---
title: "Fitcoach package workflow example"
author: "Niraj Juneja, Charles de Lassence"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fitcoach)
library(ggplot2)
```

## Example 1: Data Loader - Getting data from Fitbit API

This part explains how to connect to the Fitbit API and get your data, using `DataLoader`. 


**Step 1:** You first need to make sure that you have [registered an app](https://dev.fitbit.com/) and set it as *Personal* in order to retrieve intraday data. You will need the following credentials in order to connect the API: App name (or *OAuth 2.0 Client ID*), Client Key and Client Secret. 


**Step 2:** We initialize a new `DataLoader` object, and connect to the API with OAuth2, using the credentials described above. Note that instead of providing the credentials directly as parameters, you could point to a cache file (usually named `.httr-oauth`) using the `cache.file` parameter.

```{r eval = FALSE}
mydata <- DataLoader$new()
mydata$connect(appname = "cdlr",
               key = "227FWR",
               secret = "3089e3d1ac5dde1aa00b54a0c8661f42"
)
```


**Step 3:** We request the data and write it to JSON files using the `request` method. You need to specify the type of timeseries ('day' or 'intraday'), the list of activities ([full list here](https://dev.fitbit.com/docs/activity/)), the start and end dates, and the folder in which the JSON files will be written.

```{r eval = FALSE}
masterPath <- system.file("extdata", 
                          "daily-time-series", 
                          package = "fitcoach")

mydata$request(
    type = 'day', 
    activities = list("calories", "steps", "distance", "minutesVeryActive"), 
    start.date = "2016-01-01", 
    end.date = "2016-02-01", 
    path = masterPath
)
```

Once the JSON files have been created, they can be used for further analysis.


## Example 2: Fit Analyzer - Daily File Analysis

Examples below demonstrate usage scenarios for `FitAnalyzer`.


**Step 1:** We first need to point to a folder that contains the JSON files for "daily" file analysis. These files are created by `DataLoader.R`.

We then create a new instance of `FitAnalyzer`, passing in the folder and the goal that we want to optimize on. Goals can be the following: a) calories b) steps c) distance d) floors.

The example below uses `steps` as the goal.

```{r}
masterPath <- system.file("extdata", 
                          "daily-time-series", 
                          package = "fitcoach")

ana <- FitAnalyzer$new("steps")
```


**Step 2:** Next we get the data.frame ready for analysis. Note this data.frame is cleaned and augmented with additional data elements not present in the JSON file. E.g.: we augment `weekday`, `weekend` and mark rows that are valid.

```{r}
timeseries.frame <- 
    ana$getAnalysisFrame(folder = masterPath, 
                         analysis.type = "daily")
head(timeseries.frame)
```


**Step 3:** next we find the most important variables that are enabling meeting the goals for the person. Note this call creates a `glm` model behind the scenes and ranks the variables based on the coefficients of the glm model. You can also get the `glm` fit object to do further analysis.

```{r}
vars <- ana$findImportantVariables(tsDataFrame = timeseries.frame, 
                                   seed = 12345)
vars
```


Getting the `fit` object.

```{r}
fit <- ana$getFit()
summary(fit)
```


```{r}
par(mfrow=c(2,2))
plot(fit)
```


**Step 4:** Next, we can then plot the performance of the individual, relative to the most important variables that are making a difference. 

```{r fig.width= 7}
ana$showMostImportantCharts(tsDataFrame = timeseries.frame)
```


**Step 5:** We can also get the prediction on goal performance using the call below.

```{r}
rows.test <- timeseries.frame[sample(1:191, 1), ]
x <- createDependentVariableFrame(master = rows.test, goal = "steps")
res <- ana$predictGoal(x)
cat(paste("Prediction for the day", ": expected steps = ", round(res)))
```


*** 

## Example 3: FitAnalyzer - Intraday File Analysis

Examples below demonstrate usage scenarios for `FitAnalyzer` for **Intraday analysis**.


**Step 1**: We first need to point to a folder that contains the JSON files for *intraday* file analysis. These files are created by `DataLoader.R`.

We then create a new instance of `FitAnalyzer` passing in the folder and the goal that we want to optimize on. Goals can be the following: a) calories b) steps c) distance d) floors .

The example below uses *calories* as the goal

```{r }
masterPath <-
    system.file("extdata", "intra-daily-timeseries", package = "fitcoach")
ana <- FitAnalyzer$new("calories")
```


**Step 2:** Next we get the data.frame ready for analysis. Note that this data.frame is cleaned and augmented with additional data elements not present in the JSON file. E.g.: we augment cumulative sum during the day, weekday, weekend, etc.

```{r}
intra <- ana$getAnalysisFrame(folder = masterPath, analysis.type = "intra.day")
head(intra)
```


**Step 3:** Next we find the most important variables that are enabling meeting the goals for the person. Note: this call creates a **gbm** model behind the scenes and ranks the variables based on *relative.influence* call to `gbm` model. You can also get the `gbm` fit object to do further analysis.

```{r}
vars <- ana$findImportantVariables(intra)
vars <- sort(vars, decreasing = TRUE)
vars
```


Plot of important variables below.

```{r}
vars.frame <- data.frame(variables = names(vars), values = vars)
vars.frame$lnvalue <- log(vars.frame$values)
vars.frame <- vars.frame[1:7, ]
barplot(vars.frame$value, xlab = "variables", ylab = "relative importance", 
        names.arg = vars.frame$variables,
        cex.names = 0.65, cex.lab = 0.65, ylim = c(0.0, 0.1))
```


Summary of GBM model fit below.

```{r}
fit <- ana$getFit()
summary(fit)
```


**Step 4:** Next we can then plot the performance of the individual relative to the most important variables that are making a difference. 
For the 4 most important variables, the average value for every 15 min of a day is plotted, along with the moving average (using `geom_smooth` from `ggplot2`).

```{r fig.width= 7}
ana$showMostImportantCharts(tsDataFrame = intra)
```


**Step 5:** We can also get the prediction on goal performance using the call below.

```{r}
rows.test <- intra[sample(1:191, 1), ] # Take any random input for test
res <- ana$predictGoal(rows.test)
cat(paste("Prediction for the day", ": expected calories =", round(res)))
```


## Example 4: FitUtil - Illustration for usage of FitUtil functions

Approach to get a clean data.frame from JSON files.

```{r}
# masterPath is the folder containing JSON files
masterPath <- system.file("extdata", "daily-time-series", package = "fitcoach")

# Create the data.frame. This is not cleaned
master <- createTsMasterFrame(masterPath)

# Identify and Mark rows that are valid. i.e distance for the day >0
master <- markValidRows(master)

# Filter Valid rows only
master <- master[master$valid == TRUE, ]

# Augment data with additional information. Eg: weekday information
master <- augmentData(master)
head(master)
```

